### 2016-06-12 LITES/response.tex

\action{Section 2 was rewritten and expanded to clearly distinguish between the runtime rule (Section 2.2) and the classic analysis (Section 2.3). The runtime rule --- Equation (1) --- is stated in terms of multi-segmented deferrable tasks. }

\action{Section 4 was completely rewritten to more clearly explain that the problem of finding a suitable corresponding task set is open, and likely difficult.}

	\item 
Section 4 should be reworded. 'precisely converting a self-suspending task
system into a corresponding single-segment deferrable task set is not an
easy problem'  $\rightarrow$  'converting a self-suspending tasks system into a
single-segment deferrable task set introduced pessimism due to period
enforcement'.

\begin{response}
We do not fully understand this recommendation. In any case, we have
completely rewritten this section, which we hope that our revision addresses this point as well.
\end{response}

### 2016-06-12 locking.tex
minor revision

Binary semaphores, i.e., suspension-based locks used to realize mutually exclusive access to shared resources, are a common source of self-suspensions in multiprocessor real-time systems. When a task tries to use a resource that has already been locked, it self-suspends until the resource becomes available. Such self-suspensions due to lock contention, just like any other self-suspension, result in deferred execution and thus can detrimentally affect a task's interference on lower-priority tasks. It may thus seem natural to apply the period enforcer  to control the negative effects of blocking-induced self-suspensions.\footnote{The use of  period enforcement in combination with suspension-based locks has indeed been assumed in prior work~\cite{Raj:91}, stated as a motivation and possible use case in the original period enforcer proposal~\cite{Raj:suspension1991}, and suggested as a potential improvement elsewhere~\cite{Lak:11,LNR:09}. As we show here, this is actually unsafe.} However, as we demonstrate with two examples, it is actually unsafe to apply period enforcement to lock-induced self-suspensions.

@@

Notably, the examples in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} assume a \emph{shared-memory} locking protocol: once a lock is granted, tasks execute their own critical sections, on their assigned processors. One may wonder whether  effects similar to those described in in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} can also occur under \emph{distributed} real-time locking protocols such as the DPCP~\cite{RSL:88} or the DFLP~\cite{Br:13,Br:14}, where critical sections may be executed on dedicated \emph{synchronization processors}. In this case, the self-suspension occurs on the task's \emph{application processor}, which is different from the (remote) synchronization processor on which the critical section is executed. This separation allows employing period enforcement  only on application processors (while avoiding it on synchronization processors) without incurring the feedback cycle between blocking times and self-suspension times highlighted in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2}. However, period enforcement still invalidates all existing blocking analyses for distributed real-time semaphore protocols~\cite{RSL:88,Raj:91,Br:13} because it artificially increases blocking times if tasks contain multiple accesses to shared resources.

>>>>

In this section, we take a look at the period enforcer in the context of synchronization protocols.
Binary semaphores, i.e., suspension-based locks used to realize mutually exclusive access to shared resources, are a common source of self-suspensions in multiprocessor real-time systems. When a task tries to use a resource that has already been locked, it self-suspends until the resource becomes available. Such self-suspensions due to lock contention, just like any other self-suspension, result in deferred execution and thus can detrimentally affect a task's interference on lower-priority tasks. It may thus seem natural to apply the period enforcer  to control the negative effects of blocking-induced self-suspensions.\footnote{The use of  period enforcement in combination with suspension-based locks has indeed been assumed in prior work~\cite{Raj:91}, stated as a motivation and possible use case in the original period enforcer proposal~\cite{Raj:suspension1991}, and suggested as a potential improvement elsewhere~\cite{Lak:11,LNR:09}. As we show here, this is actually unsafe.} However, as we demonstrate with two examples, it is actually unsafe to apply period enforcement to lock-induced self-suspensions.

@@

Notably, the examples in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} assume a \emph{shared-memory} locking protocol: once a lock is granted, tasks execute their own critical sections, on their assigned processors. One may wonder whether  effects similar to those described in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2} can also occur under \emph{distributed} real-time locking protocols such as the DPCP~\cite{RSL:88} or the DFLP~\cite{Br:13,Br:14}, where critical sections may be executed on dedicated \emph{synchronization processors}. In this case, the self-suspension occurs on the task's \emph{application processor}, which is different from the (remote) synchronization processor on which the critical section is executed. This separation allows employing period enforcement  only on application processors (while avoiding it on synchronization processors) without incurring the feedback cycle between blocking times and self-suspension times highlighted in Sections~\ref{sec:sem-case-1} and~\ref{sec:sem-case-2}. However, period enforcement still invalidates all existing blocking analyses for distributed real-time semaphore protocols~\cite{RSL:88,Raj:91,Br:13} because it artificially increases blocking times if tasks contain multiple accesses to shared resources.

### 2016-06-12 transformation.tex
minor revision

To  convert a computation segment of $\tau_k$ into a single-segment deferrable task, we need to derive the segment's \emph{latest-possible arrival time}, relative to the release of a job. Formally,  for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$, we let $\rho_k^j$ denote its latest-possible arrival time, with the interpretation that, if a job of task $\tau_k$ arrives at time $t$, then  it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will arrive no later than at time $t+\rho_k^j$.

@@

However, deriving an exact bound $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ is not easy: 
even for the above ``simple'' case, Nelissen et al.'s solution~\cite{ecrts15nelissen} for calculating the exact worst-case response time requires exponential time complexity if $j \geq 2$. Furthermore, Nelissen et al. \cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that the problem of deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be very challenging indeed.\footnote{In fact, in ongoing work, this problem has recently been shown to be coNP-hard in the strong sense~\cite{Chen2016b}.}

@@

Next, we take a look at the period enforcer in the context of synchronization protocols.

>>>>

To  convert a computation segment of $\tau_k$ into a single-segment deferrable task, we need to derive the segment's \emph{latest-possible arrival time}, relative to the release of a job. Formally,  for the $j^{\mathrm{th}}$ computation segment of task $\tau_k$, we let $\rho_k^j$ denote its latest-possible arrival time, with the interpretation that, if a job of task $\tau_k$ arrives at time $t$, then  it is guaranteed that the $j^{\mathrm{th}}$ computation segment of this job will not arrive later than time $t+\rho_k^j$.

@@

However, deriving an exact bound $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ is not easy: 
even for the above ``simple'' case, Nelissen et al.'s solution~\cite{ecrts15nelissen} for calculating the exact worst-case response time requires exponential time complexity if $j \geq 2$. Furthermore, Nelissen et al. \cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that the problem of deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be very challenging indeed.\footnote{In fact, in ongoing work, verifying the schedulability of task $\tau_k$ even for this case has recently been shown to be coNP-hard in the strong sense~\cite{Chen2016b}.}

@@

### 2016-06-12 untitled.tex
minor revision

We focus exclusively on preemptive fixed-priority scheduling in this note, as the period enforcer is explicitly designed for this setting. For simplicity, we assume that tasks are indexed in order of decreasing priority (i.e., $\tau_1$ is the highest-priority task).

@@

Importantly, the tasks in $\tset$ and $\tset'$ do \emph{not} have to be of the same task model, nor does the number of tasks have to remain constant (i.e.,  $|\tset| \neq |\tset'|$ is possible). Specifically, the task set transformation underlying the analysis of the period enforcer maps each \emph{multi-}segmented self-suspending task $\tau_i \in \tset$  to $m_i$ \emph{single-}segmented self-suspending tasks in $\tset'$ (i.e., $|\tset'| = \sum_{\tset} m_i$).

@@

A segment's eligibility time is determined according to the following rule. Let $ET_{i,j}^k$ denote the eligibility time of the $k$\xth computation segment of the $j$\xth job of task $\tau_i$. Further, let $a^k_{i,j}$ denote the segment's arrival time. Finally, let $\mathit{busy}(\tau_i, t')$ denote the last time that a level-$i$ busy interval began on or prior to time $t'$ (i.e., the processor executes only $\tau_i$ or higher-priority tasks throughout the interval $[\mathit{busy}(\tau_i, t'), t']$). The period enforcer algorithm defines the segment eligibility time of the $k$\xth segment as
\begin{align}\label{eq:ET-def}
	ET_{i,j}^k & = \max\left(ET_{i,j-1}^k + T_i,\ \mathit{busy}(\tau_i, a^k_{i,j})\right),
\end{align}
where $ET_{i,0}^k = -T_i$~\cite[Section 3.1]{Raj:suspension1991}. This simple and elegant rule has the desirable effect of avoiding all back-to-back execution, which can be easily observed with an example.

@@

Figure \ref{fig:ok-with-period-enforcement} illustrates how the definition of eligibility time in Equation~\ref{eq:ET-def} restores the schedulability of the task set depicted in Figure~\ref{fig:not-ok-without-period-enforcement}. Consider the eligibility times of the second segment of task $\tau_2$.

@@

This leads to the first question: \emph{Does schedulability of the  original set of segmented self-suspending tasks (without period enforcement) imply schedulability under period enforcement?} That is, can Theorem 5 (in \cite{Raj:suspension1991}) be generalized to multi-segmented self-suspending tasks? In Section \ref{sec:unschedulable}, we answer this question in the negative.

\begin{enumerate}
	\item There exist sets of segmented self-suspending tasks that are schedulable under fixed-priority scheduling without any enforcement, but that are infeasible under period enforcement. This shows that Theorem 5 in \cite{Raj:suspension1991} has to be  used with care --- it may be applied only in the context of the transformed single-segment deferrable task set, and not in the context of the original multi-segmented self-suspending task set.
\end{enumerate}

Therefore, to apply Theorem 5 to conclude that a set of segmented self-suspending task sets remains schedulable despite period enforcement, we first have to answer the task-set transformation question: \emph{given a set of segmented self-suspending tasks $\tset$, how do we obtain a corresponding set of single-segment deferrable tasks $\tset'$ such that $\tset'$ is schedulable (without period enforcement)  only if $\tset$ is schedulable (with period enforcement)?} That is, as discussed in Section~\ref{sec:classic-analysis}, the classic analysis of the period enforcer \cite{Raj:suspension1991} presumes that it is possible to convert multi-segmented self-suspending tasks into corresponding sets of single-segment deferrable tasks, but it is left undefined in~\cite{Raj:suspension1991}  \emph{how} this central step should be accomplished. In Section \ref{sec:convert}, we make a pertinent observation.

>>>>

We focus exclusively on preemptive fixed-priority scheduling in this note, as the period enforcer is explicitly designed for this setting. For simplicity, we assume that the given tasks are indexed in a decreasing order with respect to their priorities (i.e., $\tau_1$ is the highest-priority task).

@@

Importantly, the tasks in $\tset$ and $\tset'$ do \emph{not} have to be of the same task model, nor does the number of tasks have to remain the same (i.e.,  $|\tset| \neq |\tset'|$ is possible). Specifically, the task set transformation underlying the analysis of the period enforcer maps each \emph{multi-}segmented self-suspending task $\tau_i \in \tset$  to $m_i$ \emph{single-}segmented self-suspending tasks in $\tset'$ (i.e., $|\tset'| = \sum_{\tset} m_i$).

@@

A segment's eligibility time is determined according to the following rule. Let $ET_{i,j}^k$ denote the eligibility time of the $k$\xth computation segment of the $j$\xth job of task $\tau_i$. Further, let $a^k_{i,j}$ denote the segment's arrival time. Finally, let $\mathit{busy}(\tau_i, t')$ denote the last time that a level-$i$ busy interval begins on or prior to time $t'$ (i.e., the processor executes only $\tau_i$ or higher-priority tasks throughout the interval $[\mathit{busy}(\tau_i, t'), t']$). The period enforcer algorithm defines the segment eligibility time of the $k$\xth segment as
\begin{align}\label{eq:ET-def}
	ET_{i,j}^k & = \max\left(ET_{i,j-1}^k + T_i,\ \mathit{busy}(\tau_i, a^k_{i,j})\right),
\end{align}
where $ET_{i,0}^k = -T_i$~\cite[Section 3.1]{Raj:suspension1991}. This simple and elegant rule has the desirable effect of avoiding all back-to-back execution, which can be easily observed with an example.

@@

Figure \ref{fig:ok-with-period-enforcement} illustrates how the definition of eligibility time in Equation~\eqref{eq:ET-def} restores the schedulability of the task set depicted in Figure~\ref{fig:not-ok-without-period-enforcement}. Consider the eligibility times of the second segment of task $\tau_2$.

@@

This leads to the first question: \emph{Does the schedulability of the  original set of segmented self-suspending tasks (without period enforcement) imply the schedulability under period enforcement?} That is, can Theorem 5 (in \cite{Raj:suspension1991}) be generalized to multi-segmented self-suspending tasks? In Section \ref{sec:unschedulable}, we answer this question in the negative.

\begin{enumerate}
	\item There exist sets of segmented self-suspending tasks that are schedulable under fixed-priority scheduling without any enforcement, but that are infeasible under period enforcement. This shows that Theorem 5 in \cite{Raj:suspension1991} has to be  used with care --- it may be applied only in the context of the transformed single-segment deferrable task set, but not in the context of the original multi-segmented self-suspending task set.
\end{enumerate}

Therefore, to apply Theorem 5 to conclude that a set of segmented self-suspending task sets remains schedulable despite period enforcement, we first have to answer the task-set transformation question: \emph{given a set of segmented self-suspending tasks $\tset$, how do we obtain a corresponding set of single-segment deferrable tasks $\tset'$ such that $\tset'$ is schedulable (without period enforcement)  only if $\tset$ is schedulable (with period enforcement)?} That is, as discussed in Section~\ref{sec:classic-analysis}, the classic analysis of the period enforcer \cite{Raj:suspension1991} presumes that it is possible to convert a set of multi-segmented self-suspending tasks into a corresponding set of single-segment deferrable tasks, but it is left undefined in~\cite{Raj:suspension1991}  \emph{how} this central step should be accomplished. In Section \ref{sec:convert}, we make a pertinent observation.
